---
title: "Built-in functions"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

There are a number of built-in functions in the `grainchanger` package, with their useage outlined below. While it is possible to use user-defined functions within `winmove_agg`, `nomove_agg`, and `winmove`, we welcome suggestions for additional functions. Please [add as an issue](https://github.com/laurajanegraham/grainchanger/issues) - doing it this way means we can maximise the speed of the function. 

All functions can also be used on their own, either on an object of class `winmove` or `numeric`. 

When functions are used within `winmove_agg`, `winmove`, or directly on an object of class `winmove`, they are calculated relative to within a moving window. 

When functions are used within `nomove_agg` all cells of `fine_dat` within a given cell of `coarse_dat` are aggregated using the function. 

# Current functions

```{r functions, echo = FALSE}
function_overview <- data.frame(
  `Function Name` = c("prop", "shdi", "shei", "var_range"),
  `Description` = c("Calculate the proportion of a given class", 
                    "Calculate the Shannon diversity", 
                    "Calculate the Shannon evenness", 
                    "Calculate the size of the range of values"),
  `Additional arguments` = c("lc_class (numeric)", 
                             "lc_class (numeric)",
                             "lc_class (numeric)",
                             "")
)

knitr::kable(function_overview)
```

# Shannon diversity and evenness

Shannon diversity is calculated as $$SHDI = -\sum_{i = 1}^m p_i lnp_i$$ where $p_i$ is the proportion of a given class $i$ of a total $m$ classes. 

Shannon evenness is calculated as $$SHEI = \frac{S}{ln(m)}$$

# Additional functions

We plan to add other useful functions to this small set of built-in functions, such as relevant metrics from [FRAGSTATS](https://www.umass.edu/landeco/research/fragstats/documents/fragstats.help.4.2.pdf). 

We also welcome suggestions for additional functions. Please [add as an issue](https://github.com/laurajanegraham/grainchanger/issues) - doing it this way means we can maximise the speed of the function. 