---
title: "Using Grainchanger"
author: "Laura J. Graham"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Grainchanger}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)
```

# Package overview

The primary functions of the `grainchanger` package are those which facilitate moving-window (`winmove_agg`) and direct (`nomove_agg`) data aggregation. These functions aggregate fine-grain data (`fine_dat`) to a coarse-grain (`coarse_dat`) using a function specified by the user (`agg_fun`). The moving-window method takes in an additional function (`win_fun`) which smooths the fine-grain data prior to aggregation. 

The moving-window smoothing function is also available in the package (`winmove`), as well as several [built-in functions](functions.html), and an additional utility function for use with simulated landscapes (`create_torus`).

The `winmove` function acts as a convenient wrapper to `raster::focalWeight` and `raster::focal` which takes advantage of optimised functions built into the `grainchanger` package. 

# Data aggregation & `sf`

The functions in `grainchanger` have been written to be compatible with the [`sf`](https://r-spatial.github.io/sf/) package. The online textbook [Geocomputation with R](https://geocompr.robinlovelace.net/) provides an introduction to spatial data analysis with the `sf` package. 

## Moving-window data aggregation

In this example, we show how to create a coarse grain grid over the fine-grain data using `st_make_grid` from the `sf` package. We then aggregate the package data `cat_ls` to this grid using moving-window data aggregation and plot using `ggplot2`. 

```{r}
library(grainchanger)
library(sf)
library(ggplot2)

coarse_dat <- cat_ls %>% 
  # get the bounding box
  st_bbox() %>% 
  # turn into an sfc object
  st_as_sfc() %>% 
  # negative buffer 
  st_buffer(-4) %>% 
  # make a square grid
  st_make_grid(cellsize = 19) %>% 
  # turn into sf object
  st_sf()

# we can plot this grid on top of the fine data
landscapetools::show_landscape(cat_ls) + 
  geom_sf(data = coarse_dat, alpha = 0.5)


coarse_dat$shdi_3 <- winmove_agg(coarse_dat = coarse_dat, 
                                 fine_dat = cat_ls,
                                 d = 3,
                                 type = "rectangle", 
                                 win_fun = shdi, 
                                 agg_fun = mean,
                                 is_grid = FALSE,
                                 lc_class = 1:4)

ggplot(coarse_dat, aes(fill = shdi_3)) + 
  geom_sf() + 
  theme_bw()
```

Note that when creating the grid, we made it 4 units smaller (the size of the moving window) than the fine-resolution data. The reason for this is that we avoid edge effects created when the moving window goes beyond the extent of the data. A warning is thrown if `fine_dat` is the same size as (or smaller than) `coarse_dat`. The below is an example of this, using `g_sf` from the package. 

```{r, warning = TRUE}
g_sf$shei_4 <- winmove_agg(coarse_dat = g_sf, 
                                 fine_dat = cat_ls,
                                 d = 4,
                                 type = "rectangle", 
                                 win_fun = shei, 
                                 agg_fun = mean,
                                 is_grid = FALSE,
                                 lc_class = 1:4)
```


## Direct data aggregation

In this example we show how to read in a shapefile as an `sf` object, apply direct data aggregation, and plot using `ggplot2`. `cont_ls` is a fine-resolution RasterLayer provided with the package. 

```{r}
library(sf)
library(ggplot2)

# coarse_dat <- st_read("your_file.shp")
coarse_dat <- st_read(system.file("shape/poly_sf.shp", package="grainchanger"))

coarse_dat$var_range <- nomove_agg(coarse_dat = coarse_dat,
                                   fine_dat = cont_ls,
                                   agg_fun = var_range,
                                   is_grid = FALSE)

ggplot(coarse_dat, aes(fill = var_range)) + 
  geom_sf() + 
  theme_bw()
```

# Functions

There are a number of inbuilt functions in the grainchanger package, with their useage outlined below. While it is possible to use user-defined functions within both `winmove_agg` and `nomove_agg`, we welcome suggestions for additional functions. Please [add as an issue](https://github.com/laurajanegraham/grainchanger/issues) - doing it this way means we can maximise the speed of the function. 

```{r functions, echo = FALSE}
function_overview <- data.frame(
  `Function Name` = c("prop", "shdi", "shei", "range"),
  `Description` = c("Calculate the proportion of a given class", 
                    "Calculate the Shannon diversity", 
                    "Calculate the Shannon evenness", 
                    "Calculate the range of values"),
  `Additional arguments` = c("lc_class (numeric)", 
                             "lc_class (numeric)",
                             "lc_class (numeric)",
                             "")
)

knitr::kable(function_overview)
```

# Additional utilities

## Create torus

The `create_torus` function takes as input a square or rectangular landscape and pads it by a specified radius, creating the effect of a torus. We developed this function in order to avoid edge effects when testing methods on simulated landscapes (such as those from [NLMR](https://ropensci.github.io/NLMR/)). 

```{r torus}
torus <- create_torus(cat_ls, 5)

landscapetools::show_landscape(torus)
```

## Running aggregation in parallel

The `winmove_agg` and `nomove_agg` functions have been designed to work with the [`future`](https://github.com/HenrikBengtsson/future) package. This means that with just one additional line of code, the functions can be run in parallel while taking advantage of the resources available to the user. 

```
library(future)
plan(multisession)
```
See the [`future`](https://github.com/HenrikBengtsson/future) documentation for more information. 
